### 2016-11-03 capability-machine.tex
Removed continuation relation! Expression relation now uses value relation instead (indirectly throug ref-file rel).
Changed the read write condition into a pure write condition.
Changed the conditions to use UPreds for readability. Changed some definitions to inference rules for readability.

\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage{thmtools}
\usepackage{stmaryrd}
\usepackage{natbib}
\usepackage{url}
\usepackage{array}
\usepackage{arydshln}
\usepackage{ifthen}
\usepackage{ifpdf}
\usepackage{verbatim}

@@

\newcommand{\readCond}[1]{\plainfun{readCondition}{#1}}
\newcommand{\writeCond}[1]{\plainfun{readWriteCondition}{#1}}
\newcommand{\execCond}[1]{\plainfun{executeCondition}{#1}}
\newcommand{\entryCond}[1]{\plainfun{entryCondition}{#1}}

@@

Ignoring the view part of a region, a future region is a region where the interpretation and state transitions remain the same, but the state may have changed in accordance with the state transition system: 
\begin{gather*}
  (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_1) \\
  \text{ iff } \\ 
  (s_1,s_2) \in \phi_2 \text{ and } (\phi_1,H_1) = (\phi_2,H_2)
\end{gather*}
our two future region relations both have this in common, but they also take the view into account. For the weak future world relation, the view has to stay the same:
\begin{align*}
  &\quad r_2 \futurewk r_1 \\
  &\text{ iff } \\ 
  &\quad\exists v_2,s_2,\phi_2,H_2,v_1,s_1,\phi_1,H_1\ldotp\\
  &\qquad      r_2 = (v_2,s_2,\phi_2,H_2) \text{ and } \\
  &\qquad r_1=(v_1,s_1,\phi_1,H_1) \text{ and }\\
  &\qquad v_1 = v_2 \text{ and } (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_1)\\
  &\quad \text{or} \\ 
  &\qquad r_1 = r_2 = \revoked
\end{align*}
For the strong future world relation, the $\temp$ regions are allowed to become $revoked$. Otherwise the relation is the same as the weak future world relation.
\begin{align*}
  &\quad r_2 \futurestr r_1 \\
  &\text{ iff } \\ 
  &\quad r_2 \futurewk r_1 \\
  &\quad\text{or} \\ 
  &\quad\exists s_1,\phi_1,H_1\ldotp\\
  &\qquad r_2 = \revoked \text{ and } r_1 = (\temp,s_1,\phi_1,H_1)
\end{align*}

We now get two future world relations $\futurewk$ and $\futurestr$ are defined as one would expect. That is by function extensionality, but where the regions already defined are allowed to evolve according to the appropriate future region relation:
 \begin{align*}
 W' \future_{\var{wk}/\var{str}} W & & \text{ iff } & &
   \begin{gathered}
     \dom(W') \supseteq \dom(W)\\ 
     \text{ and } \\
     \forall r \in \dom(W) \ldotp W'(r) \future_{\var{wk}/\var{str}} W(r)
   \end{gathered}
 \end{align*}

@@

\begin{align*}
    & \writeCond{} : ((\Views \rightharpoonup \Regions) \times \Globals) \fun (\Addrs^2 \times \Worlds) \monnefun \DCPred{\nats}  \\
    & \writeCond{}(\iota,\gl)(\start,\addrend,W) =  \\
    & \quad \begin{aligned}[t]
              \{ n \mid & \;\exists r \in \var{localityReg}(g,W) \ldotp \\
              & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
              & \;\qquad W(r)\nequal[n-1] \iota_{\start',\addrend'}(W(r).v) \}
            \end{aligned}
\end{align*}

\begin{align*}
    & \readCond{}(\gl)(\start,\addrend,W) =  \\
    & \quad \begin{aligned}[t]
              \{ n \mid & \;\exists r \in \var{localityReg}(g,W) \ldotp \\
              & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
              & \;\qquad W(r)\nsubeq[n-1] \iota^{\var{pwl}}_{\start',\addrend'}(W(r).v) \}
            \end{aligned}
\end{align*}
\lau{I am in doubt whether it should be $\iota^{\var{pwl}}$ in the read condition.}

\begin{align*}
  & \execCond{}(\gl)(\perm,\start,\addrend,W) = \\
  & \quad
    \begin{aligned}[t]
      \{ n \mid &  \forall n' < n \ldotp\\
      & \quad \forall W' \future W \ldotp\\
      & \qquad\forall a \in [\start,\addrend] \ldotp\\
      & \qquad \quad \npair[n']{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W')\}
    \end{aligned} \\
  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
  & \quad \text{and } \gl = \glob \Rightarrow \future = \futurestr
\end{align*}

\begin{align*}
  & \entryCond{}(\gl)(\start,\addrend,\addr,W) = \\
  & \quad
    \begin{aligned}[t]
      \{ n \mid &  \forall n' < n \ldotp\\
      & \quad \forall W' \future W \ldotp\\
      & \qquad \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')\}
    \end{aligned} \\
  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
  & \quad \text{and } \gl = \glob \Rightarrow \future = \futurestr
\end{align*}

Now define the value relation as follows:
\begin{align*}
  &\stdvr : \Worlds \monwknefun \UPred{\Words} \\
  &\stdvr\defeq \lambda \; W \ldotp 
           \begin{aligned}[t]
             & \{ \npair{i} \mid i \in \ints \} 
             \union \\
             & \{ \npair{\stdcap[(\noperm,\gl)] }  \} 
             \union \\
             & \{ \npair{\stdcap[(\readonly,\gl)] } \mid \\
             & \quad n \in \readCond{}(\gl)(\start,\addrend,W)\} 
             \union \\
             & \{ \npair{\stdcap[(\readwrite,\gl)] } \mid \\
             & \quad n \in \writeCond{}(\iota^{\var{nwl}},\gl)(\start,\addrend,W) \}
             \union \\
             & \{ \npair{\stdcap[(\readwritel,\gl)] } \mid \\
             & \quad n \in \writeCond{}(\iota^{\var{pwl}},\gl)(\start,\addrend,W) \} \union \\
             & \{ \npair{\stdcap[(\exec,\gl)]} \mid \\
             & \quad n \in \readCond{}(\gl)(\start,\addrend,W) \land \\
             & \quad n \in \execCond{}(\gl)(\exec,\start,\addrend,W) \}
             \union \\
             & \{ \npair{\stdcap[(\entry,\gl)]} \mid \\
             & \quad n \in \entryCond{}(\gl)(\start,\addrend,\addr,W)\} 
             \union \\
             & \{ \npair{\stdcap[(\rwx,\gl)]} \mid \\
             & \quad n \in \writeCond{}(\iota^{\var{nwl}},\gl)(\start,\addrend,W) \land\\
             & \quad n \in \execCond{}(\gl)(\rwx,\start,\addrend,W)  \land \\
             & \quad n \in \execCond{}(\gl)(\exec,\start,\addrend,W) \}
             \union \\
             & \{ \npair{\stdcap[(\rwlx,\gl)]} \mid \\
             & \quad n \in \writeCond{}(\iota^{\var{pwl}},\gl)(\start,\addrend,W) \land\\
             & \quad n \in \execCond{}(\gl)(\rwlx,\start,\addrend,W) \land \\
             & \quad n \in \execCond{}(\gl)(\rwx,\start,\addrend,W) \land \\
             & \quad n \in \execCond{}(\gl)(\exec,\start,\addrend,W) \}
           \end{aligned}
\end{align*}

\begin{align*}
  \observations : &  \Worlds \nefun \UPred{\Regs \times \HeapSegments} \\
  \observations (W) \defeq & \{ \npair{(\reg,\hs)} \mid \\
                           & \quad (\forall \heap_f, \heap', i \leq n \ldotp (\reg,\hs \uplus \heap_f) \step[i] (\halted,\heap')  \\
                           & \qquad \Rightarrow \exists W' \futurestr W \ldotp\exists \hs' \ldotp \heap' = \hs' \uplus \heap_f \land \heapSat[\hs']{n-i}{W'}
\end{align*}

@@

\begin{align*}
  \stdkr : &  \Worlds \monwknefun \UPred{\Words} \\
  \stdkr \defeq & \lambda W \ldotp\{ \npair{c} \mid
                     \begin{aligned}[t]
                       & \npair{c} \in \stdvr(W) \land \\
                       &  \forall W' \futurewk W, n' < n\ldotp
                       \forall \heapSat[\hs]{n'}{W'} \ldotp \forall \reg, \npair[n']{\reg} \in \stdrr(W') \ldotp \\
                       & \quad \npair[n']{(\reg\update{\pcreg}{\updatePcPerm{\var{c}}},\hs)} \in \observations(W') \}
                 \end{aligned}
\end{align*}

\begin{align*}
  \stder : & \Worlds \nefun \UPred{\Words} \\
  \stder \defeq & \lambda W \ldotp \{ \npair{\pc} \mid 
  \begin{aligned}[t]
    & \forall n' \leq n \ldotp\\
    & \quad \forall \npair[n']{\reg} \in \stdrr(W) \ldotp \\
    & \quad \quad \forall \npair[n']{c} \in \stdkr(W) \ldotp \\
    & \quad \qquad  \forall \heapSat[\hs]{n'}{W} \ldotp \\
    & \quad \qquad \quad \npair[n']{(\reg\update{r_0}{c}\update{\pcreg}{\pc},\hs)} \in \observations(W) \}
  \end{aligned}
\end{align*}

>>>>

\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage{thmtools}
\usepackage{stmaryrd}
\usepackage{natbib}
\usepackage{url}
\usepackage{array}
\usepackage{arydshln}
\usepackage{ifthen}
\usepackage{ifpdf}
\usepackage{verbatim}
\usepackage{mathpartir}

@@

\newcommand{\readCond}[1]{\plainfun{readCondition}{#1}}
\newcommand{\writeCond}[1]{\plainfun{writeCondition}{#1}}
\newcommand{\execCond}[1]{\plainfun{executeCondition}{#1}}
\newcommand{\entryCond}[1]{\plainfun{entryCondition}{#1}}

@@

Ignoring the view part of a region, a future region is a region where the interpretation and state transitions remain the same, but the state may have changed in accordance with the state transition system: 
\begin{mathpar}
  \inferrule{  (s_1,s_2) \in \phi_2 \\
               (\phi_1,H_1) = (\phi_2,H_2)}
            {  (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_1) }
\end{mathpar}
our two future region relations both have this in common, but they also take the view into account. For the weak future world relation, the view has to stay the same:
\begin{mathpar}
  \inferrule{ (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_1) \\
               v_1 = v_2 }
            { (v_2,s_2,\phi_2,H_2) \futurewk (v_1,s_1,\phi_1,H_1) }
\and
  \inferrule{ }
            { \revoked \futurewk \revoked }
\end{mathpar}
For the strong future world relation, the $\temp$ regions are allowed to become $revoked$. Otherwise the relation is the same as the weak future world relation.
\begin{mathpar}
  \inferrule{ r_2 \futurewk  r_1 }
            { r_2 \futurestr r_1 }
\and
  \inferrule{ }
            { \revoked \futurestr (\temp,s,\phi,H)}
\end{mathpar}

We now get two future world relations $\futurewk$ and $\futurestr$ are defined as one would expect. That is by function extensionality, but where the regions already defined are allowed to evolve according to the appropriate future region relation:
\begin{mathpar}
  \inferrule{ \dom(W') \supseteq \dom(W)\\ 
              \forall r \in \dom(W) \ldotp W'(r) \futurewk W(r) }
            { W' \futurestr W }
\and
  \inferrule{ \dom(W') \supseteq \dom(W)\\ 
              \forall r \in \dom(W) \ldotp W'(r) \futurestr W(r) }
            { W' \futurestr W }
\end{mathpar}

@@

\begin{align*}
    & \writeCond{} : ((\Views \rightharpoonup \Regions) \times \Globals) \fun \Worlds \monnefun \UPred{\Addrs^2}  \\
    & \writeCond{}(\iota,\gl)(W) =  \\
    & \quad \begin{aligned}[t]
              \{ \npair{(\start,\addrend)} \mid & \;\exists r \in \var{localityReg}(g,W) \ldotp \\
              & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
              & \;\qquad W(r)\nsupeq[n-1] \iota_{\start',\addrend'}(W(r).v) \}
            \end{aligned}
\end{align*}

\begin{align*}
    & \readCond{}(\gl)(W) =  \\
    & \quad \begin{aligned}[t]
              \{ \npair{(\start,\addrend)} \mid & \;\exists r \in \var{localityReg}(g,W) \ldotp \\
                        & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
                        & \;\qquad W(r)\nsubeq[n-1] \iota^{\var{pwl}}_{\start',\addrend'}(W(r).v) \}
            \end{aligned}
\end{align*}
\lau{I am in doubt whether it should be $\iota^{\var{pwl}}$ in the read condition.}

\begin{align*}
  & \execCond{}(\gl)(W) = \\
  & \quad
    \begin{aligned}[t]
      \{ \npair{(\perm,\start,\addrend)} \mid &  \forall n' < n \ldotp\\
      & \quad \forall W' \future W \ldotp\\
      & \qquad\forall a \in [\start,\addrend] \ldotp\\
      & \qquad \quad \npair[n']{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W')\}
    \end{aligned} \\
  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
  & \quad \text{and } \gl = \glob \Rightarrow \future = \futurestr
\end{align*}

\begin{align*}
  & \entryCond{}(\gl)(W) = \\
  & \quad
    \begin{aligned}[t]
      \{ \npair{(\start,\addrend,\addr)} \mid &  \forall n' < n \ldotp\\
      & \quad \forall W' \future W \ldotp\\
      & \qquad \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')\}
    \end{aligned} \\
  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
  & \quad \text{and } \gl = \glob \Rightarrow \future = \futurestr
\end{align*}

Now define the value relation as follows:
\begin{align*}
  &\stdvr : \Worlds \monwknefun \UPred{\Words} \\
  &\stdvr\defeq \lambda \; W \ldotp 
           \begin{aligned}[t]
             & \{ \npair{i} \mid i \in \ints \} 
             \union \\
             & \{ \npair{\stdcap[(\noperm,\gl)] }  \} 
             \union \\
             & \{ \npair{\stdcap[(\readonly,\gl)] } \mid \\
             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)\} 
             \union \\
             & \{ \npair{\stdcap[(\readwrite,\gl)] } \mid \\
             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^{\var{nwl}},\gl)(W) \}
             \union \\
             & \{ \npair{\stdcap[(\readwritel,\gl)] } \mid \\
             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^{\var{pwl}},\gl)(W) \}
             \union \\
             & \{ \npair{\stdcap[(\exec,\gl)]} \mid \\
             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
             & \quad \npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W) \} 
             \union \\
             & \{ \npair{\stdcap[(\entry,\gl)]} \mid \\
             & \quad \npair{(\start,\addrend,\addr)} \in \entryCond{}(\gl)(W)\} 
             \union \\
             & \{ \npair{\stdcap[(\rwx,\gl)]} \mid \\
             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^{\var{nwl}},\gl)(W) \land\\
             & \quad \npair{(\rwx,\start,\addrend)} \in \execCond{}(\gl)(W)  \land \\
             & \quad \npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W) \}
             \union \\
             & \{ \npair{\stdcap[(\rwlx,\gl)]} \mid \\
             & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
             & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^{\var{pwl}},\gl)(W) \land\\
             & \quad \npair{(\rwlx,\start,\addrend)} \in \execCond{}(\gl)(W) \land \\
             & \quad \npair{(\rwx,\start,\addrend)} \in \execCond{}(\gl)(W) \land \\
             & \quad \npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W) \}
           \end{aligned}
\end{align*}

\begin{align*}
  \observations : &  \Worlds \nefun \UPred{\Regs \times \HeapSegments} \\
  \observations (W) \defeq & \lambda W \ldotp 
                             \begin{aligned}[t]
                               \{ \npair{(\reg,\hs)} \mid & \forall \heap_f, \heap', i \leq n \ldotp (\reg,\hs \uplus \heap_f) \step[i] (\halted,\heap')  \\
                                                          & \quad \Rightarrow \exists W' \futurestr W \ldotp\exists \hs' \ldotp \heap' = \hs' \uplus \heap_f \land \heapSat[\hs']{n-i}{W'} \}
                             \end{aligned}
\end{align*}

@@

\begin{align*}
  \stder : & \Worlds \nefun \UPred{\Words} \\
  \stder \defeq & \lambda W \ldotp \{ \npair{\pc} \mid 
  \begin{aligned}[t]
    & \forall n' \leq n \ldotp\\
    & \quad \forall \npair[n']{\reg} \in \stdrr(W) \ldotp \\
    & \quad \qquad  \forall \heapSat[\hs]{n'}{W} \ldotp \\
    & \quad \qquad \quad \npair[n']{(\reg\update{\pcreg}{\pc},\hs)} \in \observations(W) \}
  \end{aligned}
\end{align*}
